Notes

### 2022-08-26 thesis/chapters/model-choice.tex
Revision up to chapter 4 complete

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Model choice, implementation and validation}\label{ch:model-choice}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
In this section we gather together several remarks on the main choices made throughout the design of our model, as well as some validation strategies that attempt to measure the goodness-of-fit of the model given the observed data.

\paragraph{The choice of \(\symbf p\).} One of the key decisions in our Bayesian modeling scheme was whether to consider the number of components \(p\) as a member of the parameter space and integrate it into the model. While theoretically we could impose a prior distribution on \(p\) as well (e.g. a categorical distribution with a fixed maximum value), we found that it would have some unwanted practical implications. For instance, it would make the implementation more complex, since the dimensionality of the parameters \(b\) and \(\tau\) would need to be fixed at a certain maximum value beforehand, but the working value of \(p\) within the MCMC algorithm would vary from one iteration to the next. In this case we would face a label switching issue \citep[c.f.][Sec.~2.3]{grollemund2019bayesian}, as we would have no way of knowing which set of parameters is ``active'' at any given time. Another possibility would be to adapt a purely Bayesian model selection technique \citep[e.g.][Ch.~7]{piironen2017comparison, gelman2013bayesian} to our framework, or even derive some model aggregation methods to combine the posterior distributions obtained for different-sized models. In the end, for the sake of simplicity we decided to let \(p\) be an hyperparameter, so that we could use any model selection criteria (e.g. BIC, DIC, cross-validation, \ldots) to select its optimal value. This is equivalent to imposing a degenerate prior distribution on \(p\). Moreover, the experiments carried out indicate that even low values of \(p\) provide sufficient flexibility in most scenarios (see Section~\ref{sec:results}). \incomment{Reversible MCMC}

@@

\paragraph{Other hyperparameters.} As for the default values of the rest of hyperparameters in~\eqref{eq:prior-linear}, several comments are in order:
\begin{itemize}
  \item For the expected value \(b_0\) we propose to use the MLE of \(b\). Although the likelihood function is rather involved, an approximation of the optimal value is enough for our purposes (see Section~\ref{sec:results} for details on the implementation). Our numerical studies suggest that the results are much better with this choice than, say, with a random or null vector.
  \item We found that the parameter \(g\) does not have as much influence on the final result, and the experimentation suggests that \(g=5\) is a good value.
  \item Lastly, we observed that the choice of \(\eta\) can have a considerable impact on the final estimator. That is why, in an effort to normalize its scale, we consider a compound parameter \(\eta = \tilde \eta \lambda_{\max}(\mathcal X_\tau'\mathcal X_\tau)\), where \(\lambda_{\max}(\mathcal X_\tau'\mathcal X_\tau)\) is the largest eigenvalue of the matrix \(\mathcal X_\tau'\mathcal X_\tau\), and \(\tilde\eta > 0\) is the actual tuning parameter. This standardization technique has been used previously in the literature; see for example \citet{grollemund2019bayesian}.
\end{itemize}

@@

@@

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Model choice, implementation and validation}\label{ch:model-choice}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

In this section we gather together several remarks on the main choices made throughout the design of our model, as well as some validation strategies that attempt to measure the goodness-of-fit of the model given the observed data.

\section{Hyperparameter selection}

\subsection*{Label switching and the choice of \(p\)}

One of the key decisions in our Bayesian modeling scheme was whether to consider the number of components \(p\) as a member of the parameter space and integrate it into the model. While theoretically we could impose a prior distribution on \(p\) as well (e.g. a categorical distribution with a fixed maximum value), we found that it would have some unwanted practical implications. For instance, it would make the implementation more complex, since the dimensionality of the parameters \(b\) and \(\tau\) would need to be fixed at a certain maximum value beforehand, but the working value of \(p\) within the MCMC algorithm would vary from one iteration to the next. In this case we would face a label switching issue \citep[c.f.][Sec.~2.3]{grollemund2019bayesian}, as we would have no way of knowing which set of parameters is ``active'' at any given time. Another possibility would be to adapt a purely Bayesian model selection technique \citep[e.g.][Ch.~7]{piironen2017comparison, gelman2013bayesian} to our framework, or even derive some model aggregation methods to combine the posterior distributions obtained for different-sized models. In the end, for the sake of simplicity we decided to let \(p\) be an hyperparameter, so that we could use any model selection criteria (e.g. BIC, DIC, cross-validation, \ldots) to select its optimal value. This is equivalent to imposing a degenerate prior distribution on \(p\). Moreover, the experiments carried out indicate that even low values of \(p\) provide sufficient flexibility in most scenarios (see Section~\ref{sec:results}). \incomment{Reversible MCMC}

@@

\subsection*{Other hyperparameters}

As for the default values of the rest of hyperparameters in~\eqref{eq:prior-linear}, several comments are in order:
\begin{itemize}
  \item For the expected value \(b_0\) we propose to use the MLE of \(b\). Although the likelihood function is rather involved, an approximation of the optimal value is enough for our purposes (see Section~\ref{sec:results} for details on the implementation). Our numerical studies suggest that the results are much better with this choice than, say, with a random or null vector.
  \item We found that the parameter \(g\) does not have as much influence on the final result, and the experimentation suggests that \(g=5\) is a good value.
  \item Lastly, we observed that the choice of \(\eta\) can have a considerable impact on the final estimator. That is why, in an effort to normalize its scale, we consider a compound parameter \(\eta = \tilde \eta \lambda_{\max}(\mathcal X_\tau'\mathcal X_\tau)\), where \(\lambda_{\max}(\mathcal X_\tau'\mathcal X_\tau)\) is the largest eigenvalue of the matrix \(\mathcal X_\tau'\mathcal X_\tau\), and \(\tilde\eta > 0\) is the actual tuning parameter. This standardization technique has been used previously in the literature; see for example \citet{grollemund2019bayesian}.
\end{itemize}

@@

\section{MCMC implementation}

@@

\section{Validation techniques}
